---
title: "animateDispersal"
author: "James Pringle"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## *Setup*

*Load libraries and setup for graphics. See the instruction files for details, in particular `03_GetData_Subset_and_Combine.html` and `04_dataStructures_and_basicPlotting.html`.*

```{r setup}
source('connectivityUtilities.R')
library(sf)
library(ggplot2)
library("rnaturalearth")
library("rnaturalearthdata")
world <- ne_countries(scale = "medium", returnclass = "sf") #get coastline data
class(world)
```

## *Get dispersal data for region.*

Use the dispersal data calculated in save_DispersalData_fromRegion.R or save_DispersalData_fromRegion.Rmd to make an animated dispersal diagram.

```{r load data from RDS}
#dataFileName='connectivityData/AprMayJune_EastFlorida'; titleStr='April-June climatological release'
#dataFileName='connectivityData/AprMayJune_QuintanaRoo'; titleStr='April-June climatological release'
dataFileName<-'connectivityData/AprMayJune_CODE_region_climatology__1m_starts'; titleStr<-'April-June CODE region climatological release'
dataFileName<-'connectivityData/AprMayJune_CODE_region_climatology__10m_fixed'; titleStr<-'April-June CODE region climatological release'
whereAllWentAllPLD<-readRDS(paste(dataFileName,'.RDS',sep=''))
whereAllStarted<-readRDS(paste(dataFileName,'_startPoints.RDS',sep=''))
```

```{r now lets animate...}
library(gganimate) #to make this work, you might also have to install the gifski library

#specify region to plot 
if (FALSE) {
  #Gulf of Mex and East Coast
  lonLimits=c(-105.0,-60) #longitude of corners
  latLimits=c(8.0,50.0) #latitude of corners
} else if (TRUE) {
  #Central California
  lonLimits=c(-126.8-4,-120.42+1) 
  latLimits=c(34.46-2,42.0+3)
}

#get initial distribution of particles

dpad=0.0
p<-ggplot(data = world) + geom_sf() + 
      coord_sf(xlim= c(lonLimits[1]-dpad, lonLimits[2]+dpad), 
               ylim = c(latLimits[1]-dpad,latLimits[2]+dpad), expand = FALSE)+
      geom_point(data = whereAllWentAllPLD, aes(x = lonTo, y = latTo,
                                                fill=log10(fracTo),
                                            colour=log10(fracTo)), 
                 size = 1, shape = 23)+
  geom_point(data=whereAllStarted,aes(x=lonFrom,y=latFrom),colour='green',alpha=0.1)+
  #gganimate specific bits
  labs(title=paste(titleStr,'PLD: {current_frame} days'))+transition_manual(PLD)

anim_save(paste(dataFileName,'.gif',sep=''),p)
print(p)    
```

If you are running this interactively, and can't find your plot in the plot window, either load the file `dispersalExpand.gif` into your browser, or type `print(p)`... Also look at any error messages that are generated -- you may need to install the `gifski` package.
